---
title: "hw5"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


problem 2 

load and describe raw dataset
```{r}
homicide_raw=read.csv("homicide.csv")
  
  
```
The homicide dataset describes homicides happened in 50 major American cities with `homicide_raw%>% nrow()` cases recorded over the past decades. There are `homicide_raw%>% ncol()` variables included in the data. Among all the variables, reported date,latitude and longitude are recorded in form of numeric values while other 9 variables are character values. 

Some race ,age and gender are unknown for victims, accounting for part of missing values other than 'na's in latitude and longitude.

For latitude and longitude, the missing amounts are `r sum(is.na(homicide_raw$lat))` and `r sum(is.na(homicide_raw$lon))` as suggested by raw homicide data

Specifically, a huge amount of victims with unknown  race, gender and age are observed in  city Albuquerque in New Mexico state. 

A typo error "Tulsa, AL" os observed, this entry could be removed in later step during data cleaning. 






data cleaning for homicide raw data
```{r}

  homicide_df=
  homicide_raw%>%janitor::clean_names()%>%
  drop_na(lat,lon)%>%
   mutate(reported_date = as.Date(as.character(reported_date), format = "%Y%m%d")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest"        ~ "unsolved",
      disposition == "Closed by arrest"      ~ "solved"
    ))%>%
       filter(city_state != "Tulsa, AL") 
```
'na's observed in latitude and longitude variables are dropped for lucidity.Report date is modified to year_month_day format for better visualization.  

In addition, a new variable 'city_state ' is created combining city and state separated by comma. Another variable disposition is created based on whether the homicide case is solved with binary outcome of 'solved' and 'unsolved'. \

The typo error "Tulsa, AL" is excluded in this data cleaning step. 




Summarize total homicide and unsolved homicide based on variable city_state
```{r}
homicide_city_df = 
  homicide_df %>% 
  select(city_state, disposition, resolution) %>% 
  group_by(city_state) %>% 
  summarize(hom_total = n(),
            hom_unsolved = sum(resolution == "unsolved"))
```

Based on the results, there are significantly much more cases in several cities. For instance, Baltimore, Chicago and Philadelphia. Amount of unsolved case is proportional to total amount of cases happened in the city . 




Using prop.test for summarizing homicide statistics in Baltimore

```{r}
baltimore = 
  prop.test(x = filter(homicide_city_df, city_state == "Baltimore, MD") %>% pull(hom_unsolved),
            n = filter(homicide_city_df, city_state == "Baltimore, MD") %>% pull(hom_total)) 

broom::tidy(baltimore) %>% 
  knitr::kable(digits = 3)


```

The estimate is 0.646 with a confidence interval of (0.628,0.663) applying 1-sample proprotion test for proportion of unsolved homicides in baltimore.


computing proportions of unsolved cases in each city using  purrr package
```{r}
test_results = 
  homicide_city_df %>% 
  mutate(prop_tests = map2(.x = hom_unsolved, .y = hom_total, ~prop.test(x = .x, n = .y)),
         tidy_tests = map(prop_tests, broom::tidy)) %>% 
  select(-prop_tests) %>% 
  unnest() %>% 
  select(city_state, estimate, conf.low, conf.high) %>% 
  mutate(city_state = fct_reorder(city_state, estimate))

test_results
```


plotting a graph displaying estimate and confidence interval for number of homicide cases in each city including error bars

```{r}
test_results %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(
    title = "Estimates and 95% CIs of Proportion of Unsolved Homicides in All Cities",
    x = "City",
    y = "Estimate")
```

Among all the cities involved, Richmond displayed the lowest  unsolved homicide proportion estimate  with estimate around 0.27 while estimate for Chicago is especially high reaching 0.75 for upper confidence level. 





problem 3  simulation


conduct simulation to conduct t test to explore power 
```{r}
# Construct function for one-sample t-test
t_test = function(n = 30, mu, sigma = 5){
  sim_data = tibble(x = rnorm(n, mean = mu, sd = sigma)) 
    
  test_data = t.test(sim_data, mu = 0, conf.level = 0.95)
  
  sim_data %>% 
    summarize(
      estimate = pull(broom::tidy(test_data), estimate),
      p_value = pull(broom::tidy(test_data), p.value)
    )
}
# Generate 5000 data from the model and repeat the t-test for mu = 0,1,2,3,4,5,6
set.seed(1)
sim_result = 
  tibble(mu = c(0:6)) %>% 
  mutate(
    output_list = map(.x = mu, ~rerun(5000, t_test(mu = .x))),
    estimate_df = map(output_list, bind_rows)
    ) %>% 
  select(-output_list) %>% 
  unnest(estimate_df)


```



proportion of null hypothesis rejected Vs true mean


```{r}
sim_result  %>% 
  group_by(mu) %>%
  filter(p_value < 0.05) %>% 
  summarize(rej_num = n(), rej_prop = rej_num/5000) %>% 
  ggplot(aes(x = mu, y = rej_prop)) + 
  geom_point(alpha = 0.8) +
  geom_line(alpha = 0.8) +
  geom_text(aes(label = round(rej_prop, 3)), vjust = -1, size = 3) + 
  scale_x_continuous(limits = c(0,6), breaks = seq(0,6,1)) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.2)) +
  labs(
    title = "rejected proprotion Vs true mean ",
    x = "True mean",
    y = "Proportion of times the null was rejected")
```

Proportion of null hypothesis rejected increases from 0.051 to 1 as true mean increases from 0 to 6. The proportion increases significantly with mu from 0 to 4  reaching 0.992 then remains as 1 from 5 to 6 as it reaches the maximum value. 



a plot including average estimated mu (y axis) Vs  true mu (x axis) 
```{r}
sim_result %>% 
  group_by(mu) %>% 
  summarise(avg_estimate = mean(estimate)) %>% 
  ggplot(aes(x = mu, y = avg_estimate)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_text(aes(label = round(avg_estimate,2)), vjust = -1, size = 3) + 
  scale_x_continuous(limits = c(0,6.5), breaks = seq(0,6,1)) +
  scale_y_continuous(limits = c(-0.1,6.5), breaks = seq(0,6,1)) +
  labs(
    title = "average estimated mean Vs true mean ",
    x = "True mean",
    y = "Average estimated mean"
  ) 

```

A linear relationship is observed between average estimated mean and true mean. Average estimated mean is positively proportional to true mean as indicated. 






Make a plot showing the average estimate of mu  on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of mu only in samples for which the null was rejected on the y axis and the true value of mu on the x axis.

```{r}


sim_rej = sim_result %>% 
  filter(p_value < 0.05) %>% 
  group_by(mu) %>% 
  summarise(avg_estimate = mean(estimate)) 
sim_result %>% 
  group_by(mu) %>% 
  summarise(avg_estimate = mean(estimate)) %>% 
  ggplot(aes(x = mu, y = avg_estimate, color = "Total samples")) +
  geom_point() +
  geom_line() + 
  geom_text(aes(label = round(avg_estimate,2)), vjust = 2, size = 5) + 
  geom_point(data = sim_rej, aes(color = "Rejected samples")) +
  geom_line(data = sim_rej, aes(x = mu, y = avg_estimate, color = "Rejected samples")) + 
  geom_text(data = sim_rej, aes(label = round(avg_estimate,2), color = "Rejected samples"), vjust = -1, size = 5) + 
  scale_x_continuous(limits = c(0,6.5), breaks = seq(0,6,1)) +
  scale_y_continuous(limits = c(-0.5,6.5), breaks = seq(0,6,1)) +
  labs(x = "true mean",
       y = "average estimate mean",
       title = "average estimates Vs true mean",
       color = "Type") +
  scale_color_manual(values = c("Total samples" = "#4393C3", "Rejected samples" = "#D6604D"))
```


Average estimated mean and true mean are similar for total samples. However, it is not the same case for rejected samples. Estimated average means fluctuate corresponding to true mean with value 1 to 2. Both estimated average means are higher than true mean and estimated average mean is even higher when true mean is 1 comapred with that when true mean is 2. 


The difference could be caused by increase in effect size. Probability of rejecting null hypothesis increases as effect size increases. Specifically, null hypothesis is more easily rejected as significantly detectable effect size is generated as true mean increases from 0 to 6. The effect size is  detected to be  significant enough to reject null hypothesis since true mean reaches 3. 
